### 2021-06-17 example.Rmd
having trouble knitting!

---
title: "Example penalized LTRC survival model with simulated data"
author: "McGough et al"
date: "6/17/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This is a brief guide to implementing a penalized survival model for left-truncated and right-censored data (LTRC), described in McGough *et al.* 2021. This vignette uses simulated data based on the Foundation Medicine and Flatiron Health Clinico-Genomic Database (CGDB), as described in the paper [repository](https://github.com/phcanalytics/coxnet-ltrc).

## Load Data and Setup

The simulated data of interest, `simdata`, contains 5000 patients and 18 variables.

```{r}
# Setup and Simulations --------------------------------------------------------

library(dplyr) 
library(glmnet)
library(ggplot2)
library(knitr)
library(rsample)
library(rmarkdown)

source("R/run_sim.R")
source("R/calibrate.R")

# Read data
sim_settings <- readRDS("sim_settings.rds")
params <- set_params(sim_settings = sim_settings, dist = "weibullPH")
simdata <- sim_survdata(params = params, n_pats = 10000)

head(simdata)
```

## Prep Model Data

Here we randomly split the simulated data, `simdata`, into train (75%) and test (25%) data. 
We also only use patients from the simulated data who are alive at entry time (`death_time` >= `entry_time`, approximately 44% of patients).
Columns 7:17 contain our predictors of interest.

```{r}
# Filter to patients alive at study entry --------------------------------------
simdata_sub <- simdata %>% filter(hidden==0)

# Split data -------------------------------------------------------------------
# Using rsample package
set.seed(7)
data_split <- initial_split(simdata_sub, prop = .75)
train_data <- training(data_split)
test_data <- testing(data_split)

# Clean up
rm(data_split)
```

## Run Model

Here we run a simple example penalized Lasso model for LTRC using `cv.glmnet()` in `glmnet` > 4.1. The lambda which minimizes the partial likelihood deviance (`lambda.min`) is chosen through 5-fold cross-validation. For more details, see the [glmnet homepage](https://glmnet.stanford.edu/articles/Coxnet.html).

```{r}
# Prep x and y matrices - train
x_train <- train_data %>% select(7:17) %>% as.matrix()
y_train <- Surv(train_data$entry_time,
          train_data$death_time,
          train_data$dead)

# Repeat for test data
x_test <- test_data %>% select(7:17) %>% as.matrix()
y_test <- Surv(test_data$entry_time,
          test_data$death_time,
          test_data$dead)

# 5-fold CV glmnet
mod <- cv.glmnet(x_train, y_train, family="cox", nfolds = 5)

```

## Model Coefficients

Choosing the best regularized model using `lambda.min`, we can assess model coefficients, which represent the log hazard ratios:
```{r}
coef(mod, s = "lambda.min")
```

Or alternatively, we may choose the model which minimizes deviance within 1 standard deviation of the minimum `lambda.1se`, which results in a sparser model:

```{r}
coef(mod, s = "lambda.1se")
```

The full path of regularization can also be plotted:
```{r}
plot(mod)
```

## Model Predictions and Performance

We will make out-of-sample predictions on the 25% hold-out set using `lambda.min`. 
```{r}
predictions <- predict(mod, x_test, s="lambda.min")
head(predictions)
```

### C-index

```{r}
get_cindex <- function(newx, newy, fit, lambda){
  survival::concordance(newy ~ predict(fit, newx = newx, s = lambda), reverse=TRUE)
}

get_cindex(x_test, y_test, mod, "lambda.min")
```

### Calibration

```{r}
# Predict survival probabilities
survfit <- survfit(mod, newx = x_test, s="lambda.min", x= x_train, y= y_train)

# Get calibration at time points
cal <- calibrate.survfit(object = survfit, times = seq(.5, 3, .5), y = y_test)

autoplot.calibrate(cal)
```